Red Wine Quality by Jeswin George
========================================================

__Aim of this project__ is to find out which chemical property influence the 
quality of red wines.
                                                                                                                            
```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(tidyr)
library(dplyr)
library(GGally)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
wine_df <- read.csv('wineQualityReds.csv', header = TRUE, sep = ',')
```

# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

Size of dataset

```{r echo=FALSE, Univariate_Plots}
dim(wine_df)
```

The dataset consists of 13 variables and 1599 observations.

Structure of dataset

```{r echo=FALSE}
wine_df$quality <- ordered(wine_df$quality, levels=c(1:10))
str(wine_df)
```

Converted the quality variable from type integer to type factor, as it denotes the ratings received by the wine.


Statistical summary of the given dataset

```{r echo=FALSE}
summary(wine_df)
```

### Analyzing variable : Quality

To check out the most frequent quality of wines available in the dataset.

```{r}
table(wine_df$quality)
```

The above stats shows the available number of different qualities of red wine 
present in the given dataset. The best quality of red wine available in the 
given dataset is ___Quality 8__  and it is totally 18 in number, Red wine with 
__Quality 7__ is 199 in number. And the most avaliable qualities of red wine are
__Quality 5__ and __QUality 6__ with 681 and 638 numbers respectively.

```{r echo=FALSE,fig.width=10}
ggplot(data = wine_df, aes(x=quality)) +
  geom_bar()+
  labs(x = 'Quality of wines', y = 'Frequency', title = 'Bargraph of frequency vs Quality of available wines')
```

So from the above two analysis: it is observed that wines with quality ratings 5 
and 6 is the most available in the given dataset. 

***

### Analyzing variable : fixed acidity

__Description of fixed acidity:__ most acids involved with wine or fixed or 
nonvolatile (do not evaporate readily).

```{r echo=FALSE,fig.width=10}
ggplot(data = wine_df, aes(x=fixed.acidity))+
  geom_histogram(binwidth = 0.5, color = 'black', fill = 'white')+
  scale_x_continuous(limits = c(4.5,16), breaks = seq(4.5,16,0.5))+
  labs(title = 'Histogram of fixed acidity')
```

From the above diagram it is clear that most of the wines in the given dataset 
have fixed acidity between 6.5 and 8 g/dm^3^.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=10}
ggplot(data = wine_df, aes(x=fixed.acidity))+
  geom_histogram(binwidth = 1, color = 'black', fill = 'white')+
  scale_x_continuous(limits = c(4.5,16), breaks = seq(4.5,16,1))+
  facet_wrap(  ~ quality, ncol=2, scales = 'free_y')+
  labs(title = 'Faceting the distribution of fixed.acidity based on wine quality')
```

The above facetted graph, confirms the previous two analysis that the red wines 
with __Quality 5 and 6__ are the most available in the dataset and their fixed 
acidity mostly ranges between 6.5 g/dm^3^ and 8 g/dm^3^ respectively. And the best
two  available qualities of wine __Quality 7 and 8__ have a very low count and 
their  value of fixed acidity usually varies between 6.5 and 11.5 g/dm^3^. 
Also __Quality 3 and 4__ wines have value of fixed acidity varying between 6,5 
and 8.5 g/dm^3^.


```{r echo=FALSE}
by(wine_df$fixed.acidity, wine_df$quality, summary)
```

The above statistics show the summary of fixed acidity present in different 
qualities of red wine.

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=10}
ggplot(data = wine_df, aes(x=quality, y=fixed.acidity))+
  geom_boxplot()+
  coord_cartesian(ylim = c(6,10.5))+
  labs(title = 'Boxplot of fixed acidity vs Quality of red wine')
```

The __Quality 8__ of red wine has median fixed acidity  of 8.25 g/dm^3^ and mean
fixed acidity of 8.567 g/dm^3^. The __Quality 7__ of red wine has median fixed 
acidity of 8.8 g/dm^3^ and mean fixed acidity of 8.872 g/dm^3^.  

The most available __Quality 5__ of red wine has median fixed-acidity of 7.8 
g/dm^3^ and mean fixed-acidity of 8.167 g/dm^3^. And __Quality 6__ of red wine  
has median fixed-acidity of 7.9 g/dm^3^ and mean fixed-acidity of 8.347 g/dm^3^.

And the available poor qualities of red wine namely __Quality 3 and Quality 4__
have median fixed acidity of 7.5 g/dm^3^ each and their mean value of fixed 
acidity is 8.36 and 7.79 g/dm^3^ respectively.

So from the above anlaysis, fixed acidity volume does somewhat improve the 
quality of the condidered red wine.

***

### Analyzing variable : volatile acidity

__Description of volatile acidity:__ the amount of acetic acid in wine, which at
too high of levels can lead to an unpleasant, vinegar taste

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=10}
ggplot(data = wine_df, aes(x=volatile.acidity))+
  geom_histogram(binwidth = 0.1, color = 'black', fill = 'white')+
  scale_x_continuous(limits = c(0,1.6), breaks = seq(0,1.6,0.1))+
  labs(title = 'Histogram of volatile acidity content in the available red wines')
```

From the above histogram of volatile acidity - most of the wines have volatile 
acidity in the range of 0.3 to 0.7 g/dm^3^.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=10}
ggplot(data = wine_df, aes(x=volatile.acidity))+
  geom_histogram(binwidth = 0.1, color = 'black', fill = 'white')+
  scale_x_continuous(limits = c(0,1.6), breaks = seq(0,1.6,0.1))+
  facet_wrap(  ~ quality, ncol=2, scales = 'free_y')+
  labs(title = 'Faceting the distribution of volatile.acidity based on wine quality')
```

Most of the red wines with __Quality 8 and Quality 7__ have volatile acidity 
usually  ranging from 0.2 to 0.4 g/dm^3^. Also their distribution is 
right-skewed,so we can infer from this that good quality red wines have usually
low values of volatile acidity.

Red wines of __Quality 6 and 5__ are roughly normally distributed and their 
values usually range from 0.2 to 0.7 g/dm^3^.

Poor quality red wines of __Quality 3 and 4__ have their values ranging from
0.4 to 1 g/dm^3^.

```{r echo=FALSE}
by(wine_df$volatile.acidity, wine_df$quality, summary)
```

The above statistics show the summary of volatile acidity present in different 
qualities of red wine.

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=10}
ggplot(data = wine_df, aes(x=quality, y = volatile.acidity))+
  geom_boxplot()+
  coord_cartesian(ylim = c(0.25,1.05))+
  labs(title = 'Boxplot of volatile acidity vs Quality of red wine')
```

The above box plot shows that good quality red wines have lower amounts of 
volatile acids present in them.

The best available red wine of __Quality 8 and Quality 7__ have median volatile 
acidity of 0.37 g/dm^3^ each and mean volatile acidity of 0.4233 and 0.4033 
g/dm^3^. And as the quality of red wine degrades amount of volatile acidity 
present in them also increases.

***

### Analyzing variable : Citric Acid

__Description of Citric acid:__  found in small quantities, citric acid can add 
'freshness' and flavor to wines.

```{r echo=FALSE,fig.width=10}
ggplot(data = wine_df, aes(x=citric.acid))+
  geom_histogram(binwidth = 0.05, color = 'black', fill = 'white')+
  scale_x_continuous(limits = c(0,1), breaks = seq(0,1,0.1))+
  labs(title = 'Histogram of Citric acid content in the available red wines')
```

The above histogram on amount of citric acid in the red wines is right-skewed.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=10}
ggplot(data = wine_df, aes(x=citric.acid))+
  geom_histogram(binwidth = 0.05, color = 'black', fill = 'white')+
  scale_x_continuous(limits = c(0,1), breaks = seq(0,1,0.1))+
  facet_wrap(  ~ quality, ncol=2, scales = 'free_y')+
  labs(title = 'Faceting the distribution of citric acid based on wine quality')
```

The above faceted graph shows the amount of citric acid present in different 
qualities of red wine.  
Good red wines of __Quality 8 and 7__ have citric acid in the range 0.3 to 0.55 
g/dm^3^.  
Red wines of __Quality 6, 5 and 4__ have  roughly right skewed distribution and the
amount of citric acid in them range from 0 t0 0.6 g/dm^3^  and their values
usually lie around 0.35 to 0.6 g/dm^3^.   
Poor quality red wines of __Quality 3__ use citric acid in the range from 
0 to 0.6 g/dm^3^.

```{r echo=FALSE}
by(wine_df$citric.acid, wine_df$quality, summary)
```

The above statistics show the summary of citric acid present in different 
qualities of red wine.

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=10}
ggplot(data = wine_df, aes(x=quality, y = citric.acid))+
  geom_boxplot()+
  coord_cartesian(ylim = c(0,0.55))+
  labs(title = 'Boxplot of citric acid vs Quality of red wine')
```

The above box plot shows that good quality red wines have higher amounts of 
citric acids present in them.

The best available red wine of __Quality 8 and Quality 7__ have median citric 
acid of 0.42 g/dm^3^ and 0.4 g/dm^3^ and mean volatile acidity of 0.3911 g/dm^3^
and 0.3752 g/dm^3^. And as the quality of red wine degrades amount of citric 
acid present in them also deceases.

***

### Analyzing variable : residual sugar

__Description of residual sugar:__ the amount of sugar remaining after 
fermentation stops, it is rare to find wines with less than 1 gram/liter and 
wines with greater than 45 grams/liter are considered sweet.

```{r echo=FALSE, fig.width=10}
ggplot(wine_df, aes(x=residual.sugar))+
  geom_histogram(binwidth = 0.5)+
  scale_x_continuous(limits = c(0.5,16), breaks = seq(0.5,16,1))+
  labs(title = 'Histogram of residual sugar content in the available red wines')
```

The distribution obtained above is right-skewed distribution. And most of the 
available red wines have residual sugar in the range of 1.5 to 2.5 grams/liter.

```{r echo=FALSE, fig.width=10}
ggplot(wine_df, aes(x=residual.sugar))+
  geom_histogram(binwidth = 0.5)+
  scale_x_continuous(limits = c(0.5,16), breaks = seq(0.5,16,1))+
  facet_wrap( ~ quality,ncol=2, scales = 'free_y')+
  labs(title = 'Faceting the distribution of residual sugar content based on wine quality in the available red wines')
```

The above diagram shows that almost all qualities of red wines use residual 
sugar in amounts between 1.5 to 2.5 grams/liter. And scarcely do the values of 
the residual sugar level varies. So chances are that the presence of residual 
sugar must be almost fixed in all red wines.

```{r echo=FALSE}
by(wine_df$residual.sugar, wine_df$quality,summary)
```

The above statistics show the summary of residual sugar present in different 
qualities of red wine.

```{r echo=FALSE, fig.width=10}
ggplot(wine_df, aes(x = quality, y = residual.sugar))+
  geom_boxplot()+
  coord_cartesian(ylim = c(1.5,3.5))+
  labs(title = 'Boxplot of amount of residual sugar level vs Quality of red wines')
```

The above graph using boxplots and the statistical summary above that proves 
that in almost all qualities of red wines the median quantity of residual sugar
ranges from 2.1 to 2.3 grams/liter. And the mean quantity of residual sugar is 
around  2.5 grams/liter.

***

### Analyzing variable : chlorides

__Description of chlorides:__ The amount of salt in the wine.

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=10}
ggplot(wine_df, aes(x=chlorides))+
  geom_histogram(binwidth = 0.025,color = 'black', fill='white')+
  scale_x_continuous(limits = c(0,0.62), breaks = seq(0,0.62,0.05))+
  labs(title = 'Histogram of amount of chlorides in the available red wines')
```

The above histogram has a right skewed distribution. And it can be concluded 
from the distribution that the chlorides added to the red wines are in the range
of 0.05 to 0.1 grams.

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=10}
ggplot(wine_df, aes(x=chlorides))+
  geom_histogram(binwidth = 0.025,color = 'black', fill='white')+
  scale_x_continuous(limits = c(0,0.62), breaks = seq(0,0.62,0.05))+
  facet_wrap( ~ quality, ncol=2, scales = 'free_y')+
  labs(title = 'Faceting the distribution of chlorides based on wine quality in the available red wines')
```

The above faceted graph shows almost all qualities of wine have their chlorides 
amount in the range of 0.05 to 1 grams. So it can be inferred that it is 
necessary to have chlorides in the abive mentioned range for the preapartaion of
the above wine.

```{r}
by(wine_df$chlorides, wine_df$quality, summary)
```

```{r}
ggplot(wine_df, aes(x=quality,y=chlorides))+
  geom_boxplot()+
  coord_cartesian(ylim = c(0.05,0.15))+
  labs(title='Boxplot of amount of chlorides vs the quality of red wine')
```

Using the boxplot and statistical summary it is seen that chlorides are used in 
small quanities of 0.05 to 0.15 grams. And the best available red wine of 
__quality 8__ uses the least median amount of 0.0705 grams and as the quality of
red wine decreases the chloride content increases along with it.   

So it can be assumed that the best quality of red wine has the least chloride
content in it.

***

### Analyzing variable : free.sulfur.dioxide

__Description of free.sulfur.dioxide:__ The free form of SO2 exists in 
equilibrium between molecular SO2 (as a dissolved gas) and bisulfite ion; it 
prevents microbial growth and the oxidation of wine

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=10}
ggplot(wine_df, aes(x=free.sulfur.dioxide))+
  geom_histogram(binwidth = 1, color = 'black', fill='white')+
  scale_x_continuous(limits = c(1,72), breaks = seq(1,72,1))+
  labs(title = 'Histogram of free.sulfur.dioxide content in the available red wines')
```

The distribution of free.sulfur.dioxide content in the available red wines is 
right-skewed in nature and its mode value at 5.

```{r echo=FALSE, message=FALSE, warning=FALSE,fig.width=10}
ggplot(wine_df, aes(x=free.sulfur.dioxide))+
  geom_histogram(binwidth = 1, color = 'black', fill='white')+
  scale_x_continuous(limits = c(1,72), breaks = seq(1,72,1))+
  facet_wrap( ~ quality, ncol = 2) +
  labs(title = 'Faceting the distribution of free.sulfur.dioxide based on wine quality')
```

The faceted distribution of free suplfur dioxide based on quality shows that 

























Faceting the distribution of chlorides based on wine quality in the available red wines

echo=FALSE, message=FALSE, warning=FALSE,fig.width=10





































> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!